\chapter*{Acknowledgements}
%\addcontentsline{toc}{chapter}{Agradecimentos}

In embarking on this thesis development journey, I am grateful for the support and guidance provided by many individuals and institutions. Their contributions played pivotal roles in shaping the trajectory of this work.

I am grateful to my distinguished thesis supervisors, Luís Paulo Reis and Gilberto Bernardes. Their dedicated mentorship and guidance have been pivotal in guiding this research. Their expertise and encouragement have enlightened every step of this journey, and for that, I am profoundly appreciative.

The Artificial Intelligence and Computer Science Laboratory (LIACC) deserves thanks for the necessary server infrastructure for my project's development. 

The academic atmosphere at the Department of Informatics and FEUP (Faculdade de Engenharia da Universidade do Porto) contributed to my intellectual curiosity and academic growth. The institution's commitment to excellence supported my development as a student.

I am grateful for the support of my family, as well as the contributions of individuals, institutions, and communities involved in this endeavor. To my mother, father, and brother, Fábio, I appreciate their support. I pay tribute to my beloved cat, Kikka, whose presence has brought joy to my life and serves as an inspiration for my work. I am grateful to my close friends for their enduring support throughout this journey.

I humbly acknowledge the larger AI community as a critical driving force behind technological advancement. The collaborative effort in this community has been the primary driving force behind the advancements that will propel us into the future.

I appreciate the music world, as it has been a crucial aspect of my identity, in addition to my pursuit of computer science. Music stimulates emotions, demonstrates complex patterns, and shapes our perception of the world, which has motivated me during my journey. Being a musician, I find parallels between harmonies in melodies and algorithms. Furthermore, I express gratitude to the institutions and communities that uphold musical expression. Their dedication to the art of sound has enriched my own creative ventures and surely influenced how I explore sound in this thesis.

This journey demonstrates the effectiveness of collaboration, shared contributions, and the synergistic relationship between disciplines. This work testifies to our joint pursuit of knowledge and innovation, and I am grateful to have contributed to this exceptional and harmonious effort.

\vspace{10mm}
\flushleft{Márcio Duarte}
