\section{Generative Model Development} \label{sec:sol-models}

Audio synthesis is the process of creating artificial sounds from scratch. It is a challenging task that requires a deep understanding of the nature and structure of sound, as well as the ability to generate realistic and diverse audio samples. Generative models have been widely used in image synthesis, but their application to audio synthesis is relatively new and less explored.

This section presents the development of a novel generative model for audio synthesis, called GANmix, which combines elements of both a \ac{GAN} and a \ac{VAE}. GANmix exploits the strengths of both models to produce high-quality and diverse audio samples in a computationally efficient manner. GANmix also allows for latent space manipulation, which can be used to control various aspects of the generated audio, such as pitch, timbre, and style.

Before introducing GANmix, this section also describes a series of exploratory experiments that provide a valuable learning phase for the development of the main model. The experiments cover different aspects of audio synthesis, such as audio representation, model architecture, training process, dataset, and evaluation metrics. The experiments aim to provide insight into the performance, limitations, and potential of different generative audio models.

The section is organized as follows: Section~\ref{sec:findings} presents the exploratory experiments that inform the development of GANmix. Section~\ref{sec:ganmix} introduces GANmix as a novel technique for audio synthesis. 

\input{src/chapters/4-solution/models/findings}
\input{src/chapters/4-solution/models/ganmix}