\chapter{Overview of the State-of-the-Art Techniques} \label{chap:sota}

\minitoc

This chapter provides an objective review of the pertinent literature on \ac{DL} and audio processing. Its goal is to offer readers a comprehensive understanding of the current state of knowledge in the field and its evolution. Recent advancements in tools for sound comprehension and generation are enhancing existing models, enabling the modeling of new sounds more efficiently and rapidly~\cite{tahiroglu_-terity_2020}. Specifically, this chapter plays the following essential roles:

\begin{enumerate}
    \item To establish the context for the research problem: By reviewing the existing literature, the chapter sets the stage for the research problem and provides a basis for understanding its significance and importance.
    \item To identify gaps in the literature: The chapter helps to identify areas where further research is needed, as well as potential opportunities for contribution.
    \item To provide a foundation for the research design: The chapter helps to inform the design of the research study by highlighting previous research and its limitations.
    \item To demonstrate the originality of the research: By reviewing the existing literature, the chapter helps demonstrate the originality of the research problem and the thesis's contribution to the field.
    \item To position the thesis within the larger context of the field: The chapter helps to position the thesis within the larger context of the field, demonstrating its relevance and significance.
\end{enumerate}

The chapter is divided into two main sections: Background and Related Work. The Background section, present in \ref{sec:background}, discusses previous work that is important for understanding the context of the research problem, even though it does not address the same problem as the thesis. On the other hand, the Related Work section, present in \ref{sec:related-work}, focuses on issues that are similar or closely related to the research problem addressed in the thesis. This chapter serves as a foundation for the research problem and contributes to the overall contribution of the thesis.

\input{src/chapters/2-state-of-the-art/background}
\input{src/chapters/2-state-of-the-art/related-work}