\chapter{Development and Implementation} \label{chap:solution}

\minitoc

This thesis's development and implementation chapter corresponds to the crucial task of developing the solution. This chapter comprehensively explores the methodology, dataset analysis, generative model development, and research plan. This chapter aims to provide a detailed account of the steps taken to achieve the research objectives and contribute to the field of generative AI models for audio.

In Section~\ref{sec:sol-approach}, the author outlines the approach taken in developing this thesis. Section~\ref{sec:sol-datasets} details the various datasets considered in this research. A thorough description and analysis of these datasets is provided, highlighting their characteristics and potential challenges. 

Section~\ref{sec:sol-models} forms the core of this research. There, the author presents the development of several generative models, including the central model, GANmix. The model architecture, training process, and evaluation are discussed, demonstrating their contributions to the field and their potential for generating audio samples.

In addition, Section~\ref{sec:work-plan} outlines the timeline for the development of this thesis. A roadmap of key milestones and stages is provided.

Finally, practical information and resources, such as the source code, are available at \url{https://ctrlmarcio.github.io/audio-gen-ai/}. This allows interested readers to access and explore the codebase, facilitating further research and collaboration.

\input{src/chapters/4-solution/approach.tex}
\input{src/chapters/4-solution/datasets.tex}
\input{src/chapters/4-solution/models.tex}
\input{src/chapters/4-solution/work-plan.tex}