\chapter{Evaluation and Discussion} \label{chap:results}

\minitoc

The assessment of generative \ac{AI} models for audio is pivotal in advancing the field of audio generation and synthesis. It is imperative to comprehend the capabilities and limitations of these models for their effective development and implementation. This chapter presents the outcomes and analysis of experiments performed with the GANmix architecture, utilizing objective metrics based on loss functions. The aim of this study is to assess the efficacy of generative \ac{AI} models in producing audio, and to draw insights from their performance.

The experiments were conducted utilizing different hardware configurations, For \acp{GPU}, it consisted of Kaggle's Tesla V100, \ac{LIACC}'s GeForce GTX 1080, and \ac{LIACC}'s GPU Quadro RTX 8000. These configurations, along with the utilization of the PyTorch deep learning framework, allowed for proficient training and assessment of the generative \ac{AI} models.

In this chapter, the author describes the experimental setup in Section~\ref{sec:res-setup}, comprising both hardware and software configurations, alongside the GANmix architecture. The experiments' outcomes are presented in Section~\ref{sec:res-presentation} with line plots exhibiting the generator and discriminator's loss evolution per epoch. Additionally, the resulting spectrograms are presented to offer a graphical depiction of the produced audio.

Following the presentation of results, the author analyzes and interprets the findings in Section~\ref{sec:res-analysis}, identifying trends, patterns, and significant insights that emerged from the experiments. Comparisons are conducted among various models and variations in the GANmix framework for evaluating their respective performances.

Furthermore, in Section~\ref{sec:limitations}, the author addresses any limitations or challenges encountered during the evaluation process. These factors may include hardware limitations or other constraints that impact the performance of the models.

\input{src/chapters/4.5-results/setup}
\input{src/chapters/4.5-results/presentation}
\input{src/chapters/4.5-results/analysis}
\input{src/chapters/4.5-results/limitations}