\section{Presentation of Results} \label{sec:res-presentation}

This section details the outcomes of experiments that aimed to develop and evaluate generative \ac{AI} models for audio production. The chief objectives of these experiments were to explore the potential of these models to generate audio with exceptional quality and to assess their performance in various circumstances. The experiments addressed specific research questions and hypotheses. 

This section presents a thorough summary of the experiments conducted in this research. Each experiment was intentionally designed to evaluate and test the effectiveness of the generated \ac{AI} models for audio.

Each experiment begins with an introduction that highlights the motivation and specific objectives, thereby establishing the context and rationale for the experiment. In addition, the text provides a detailed description of the hardware, software, and architecture used in each experiment to ensure transparency and reproducibility.

The performance of the generative \ac{AI} models is assessed by presenting line plots showing the evolving losses throughout the training process. In addition, spectrograms are displayed to visually represent the generated audio, allowing a qualitative evaluation of the model's performance.

By systematically organizing the experimental details, a thorough understanding of the experimental procedure and results is established. This section provides the basis for analyzing and discussing the effectiveness of the generative \ac{AI} models. These results can also be seen in table format in Appendix~\ref{ann:ganmix-results}.

\input{src/chapters/4.5-results/presentation/exp1}
\input{src/chapters/4.5-results/presentation/exp2}
\input{src/chapters/4.5-results/presentation/exp3}
\input{src/chapters/4.5-results/presentation/exp4}
\input{src/chapters/4.5-results/presentation/exp5}
\input{src/chapters/4.5-results/presentation/exp6}
\input{src/chapters/4.5-results/presentation/exp7}
\input{src/chapters/4.5-results/presentation/exp8}
\input{src/chapters/4.5-results/presentation/exp9}
\input{src/chapters/4.5-results/presentation/exp10}