\subsection{Foundations for Enhancing Generative Models for Audio} \label{sec:parallel-tasks}

To develop generative models for audio, it is necessary to address several factors that impact their performance and quality. This thesis concentrates on three main areas: data augmentation, evaluation metrics, and data embedding.

\textit{Data augmentation} is the process of applying transformations to the original data to increase its size and diversity. This can help overcome the limitations of small or imbalanced datasets and improve the generalization ability of generative models. Different types of data augmentation techniques for sound generation and their effects on the model outcomes are discussed in Sextion \ref{sec:data-augmentation}.

\textit{Evaluation metrics} are the methods used to measure the quality and diversity of the generated sounds. They provide a way to compare different generative models and assess their strengths and weaknesses. However, evaluating sound generation is not trivial, as it involves objective and subjective criteria. We review various evaluation metrics for sound generation and their advantages and disadvantages in Section \ref{sec:evaluation}

\textit{Data embedding} is the technique of converting data into numerical representations that capture its essential features and characteristics. This can facilitate the learning process of generative models and enhance their expressiveness and efficiency. We explore different data embedding methods in section \ref{sec:text-embedding}.

\input{src/chapters/2-state-of-the-art/background/more/data-augmentation}
\input{src/chapters/2-state-of-the-art/background/more/evaluation}
\input{src/chapters/2-state-of-the-art/background/more/embedding}