\section*{Abstract Notes}

\begin{itemize}
\item 1500-2000 characters
    \item only one paragraph
    \item The abstract should begin with a brief but precise statement of the problem or issue, followed by a description of the research method and design, the major findings, and the conclusions reached.
\end{itemize}

abstract parts:

\begin{enumerate}
    \item introduction
    \item methods
    \item results
    \item discussion
\end{enumerate}

\textbf{Introduction}

\begin{itemize}
    \item brief context (w/o details)
    \item if the abstract has scientific terms, define them
    \item After identifying the problem, state the objective of your research. Use verbs like “investigate,” “test,” “analyze,” or “evaluate” to describe exactly what you set out to do
    \item present or past simple tense, no future
\end{itemize}

\textbf{Methods}

Next, indicate the research methods that you used to answer your question. This part should be a straightforward description of what you did in one or two sentences. It is usually written in the past simple tense, as it refers to completed actions.

Don't evaluate validity or obstacles here—the goal is not to give an account of the methodology's strengths and weaknesses, but to give the reader a quick insight into the overall approach and procedures you used.

\textbf{Results}

Depending on how long and complex your research is, you may not be able to include all results here. Try to highlight only the most important findings that will allow the reader to understand your conclusions.

\textbf{Discussion}

\begin{itemize}
    \item present simple tense
    \item if there are important limitations to research (like sample size), mention them here
    \item If your aim was to solve a practical problem, your discussion might include recommendations for implementation. If relevant, you can briefly make suggestions for further research.
\end{itemize}

\section*{Abstract}

Nowadays, audio is at the heart of most generated content. Terabytes of audio are published every day. Usually, manual labor is behind this enormous amount of generated audio. If some creator desires a specific sound, they must research it under online databases, synthesize it, or even go outside and record it. This amount of work could be a barrier to content creation, especially if such sounds are complex or too specific. This dissertation studies state-of-the-art generative AI models and follows the implementation of a system capable of generating sound snippets with a high degree of fidelity using such models. A generated sound snippet will have high fidelity if it is difficult for a human to differentiate it from a real-world sound. Audio files only present one dimension, i.e., the amplitude of the sound wave at a given moment. Models capable of generating images with outstanding results are becoming mainstream. Since audio files are theoretically simpler than image files, as they only present one data dimension, opposing three-dimensional images, these audio generative models should successfully generate believable sounds. This work should accelerate the production process for content creators, sound engineers, and everyone interested in creating audio outputs.