\chapter{Introduction} \label{chap:intro}

\minitoc

Audio is essential for shaping human experiences and interactions in the digital age. Audio content, ranging from podcasts and music to sound effects and immersive environments, is pervasive in our lives and enhances multimedia experiences. However, generating high-quality, diverse, and contextually relevant audio is still a time-consuming and labor-intensive process. As the demand for audio content continues to increase, there is a need for innovative solutions that can simplify the process and enable creators to generate various audio content with less effort.

Recent advances in \ac{AI} and \ac{DL} are revolutionizing different domains, such as image generation and text-to-speech synthesis. These cutting-edge models exhibit exceptional abilities to generate high-quality content from basic textual inputs. This thesis is motivated by these successful models and explores the potential of \ac{AI}-driven generative models to synthesize audio snippets based on textual descriptions. This thesis is not constrained to specific domains, like music or speech, but instead studies and implements systems suitable for a wide range of audio prompts.

The motivation behind this work lies in the potential benefits that an end-to-end audio generative model can offer content creators, sound engineers, and other stakeholders in the audio production ecosystem. By reducing the time and effort required to generate unique and high-quality audio samples, this research aims to contribute to the democratization of audio content creation and facilitate new avenues for creative expression.

This chapter provides an introduction to the research study and sets the stage for the exploration of the topic. It aims to supply a comprehensive overview of the present study's context, motivation, objectives, and structure.

The context information can be found in section \ref{sec:context}. Its focus is to provide the necessary background for the research study and to explain its importance. Section \ref{sec:motivation} explores the motivation behind the study. The purpose of this section is to explain the necessity of the research and the knowledge gap it aims to address. Next, in section \ref{sec:objectives}, the research study's objectives are described. This section aims to explain the specific objectives and outcomes the research seeks to accomplish. The structure of the dissertation is presented in Section \ref{sec:struct}. This section offers an organized overview of the remaining chapters.

\input{src/chapters/1-introduction/context}
\input{src/chapters/1-introduction/motivation}
\input{src/chapters/1-introduction/objectives}
\input{src/chapters/1-introduction/struct}
