\chapter{The Synthesis Problem}\label{chap:problem}

\minitoc

This chapter explores the intricate domain of sound synthesis through advanced generative \ac{AI} models that operate on textual input. The exploration seeks to unravel the complexities and nuances inherent in the fusion of \ac{AI} and sound engineering, resulting in exceptional fidelity and realism in auditory creations.

Sound synthesis, a venerable discipline in the aural arts, stands at the intersection of traditional craftsmanship and contemporary technology. The digital age is ushering in a profound transformation in which the marriage of textual prompts and generative \ac{AI} creates a symphony of possibilities for creating immersive auditory experiences.

Generative \ac{AI} orchestrates a paradigm shift in the landscape of sound synthesis. This chapter seeks to illuminate the nascent symphony of generative \ac{AI}'s role in creating intricate soundscapes, shedding light on its potential to revolutionize content creation, music composition, and multimedia production.

This chapter delves into the synthesis problem through several key sections. The first Section (\ref{sec:problem-definition}) defines the core challenges. Then, Section~\ref{sec:significance} highlights the broader implications. Section~\ref{sec:prob-datasets} examines the underlying data. Section~\ref{sec:limitations} outlines the limitations of the research. Model fine-tuning is explored in Section~\ref{sec:parametric-control}, while architectural subtleties are revealed in Section~\ref{sec:model}.


\input{src/chapters/3-problem/definition}
\input{src/chapters/3-problem/applications}
\input{src/chapters/3-problem/datasets}
\input{src/chapters/3-problem/hardware-constraints}
\input{src/chapters/3-problem/parametric-control}
\input{src/chapters/3-problem/model}