\subsection{Deep Learning}\label{sec:deep-learning}

As previously mentioned, the growth of \acf{DL} began in the early 2000s as a response to the challenge of handling vast quantities of data. Fundamentally, \ac{DL} stems from the application of \ac{ML} to process large amounts of data. To be precise, \ac{DL} is a subfield of \ac{ML} that employs multiple levels of information processing and abstraction to learn and represent features, as demonstrated by Deng et al. \cite{deng_deep_2014}. Subsequently, the extracted features can be utilized for classification, regression, and other modeling techniques. In the past, such features were manually selected by humans.

This study uses \ac{DL} for sound generation because it offers several advantages over traditional sound generation techniques. By being data-driven, these models can generate new sounds based on sounds it has heard before. On traditional methods, these sounds would have to come from the inspiration of their human creator. Besides, end-to-end generation, from text to sound generation, is only possible through \ac{DL}. The model has to extract features from the text, learn features from thousands or millions of sounds, and correlate both. This highly complex task can only be achieved with \ac{DL} techniques.

This section presents traditional \ac{DL} architectures and their evolution to generative \ac{DL} architectures.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{src/chapters/2-state-of-the-art/background/deep-learning-models/deep-learning-architectures.tex}
\input{src/chapters/2-state-of-the-art/background/deep-learning-models/dl-basics}
\input{src/chapters/2-state-of-the-art/background/deep-learning-models/generative-deep-learning-architectures.tex}