\section{Background} \label{sec:background}

This dissertation is situated in a vast body of research on \ac{DL}. This section positions the current dissertation within this context. This work is not intended to explain other technologies that solve similar problems, but rather to explain other technologies that are useful in addressing the problem at hand.

The dissertation aims at readers with basic \ac{ML} knowledge, but not necessarily basics on \ac{DL}. No basis on sound besides 8th-grade physics is needed, also.

The dissertation will start by explaining how sound can be processed digitally in subsection \ref{sec:sound}, keeping this objective in mind. Then, in subsection \ref{sec:deep-learning}, it will provide a comprehensive review of general deep learning architectures and techniques, emphasizing their importance in sound synthesis. In subsection \ref{sec:parallel-tasks}, the subsequent part will concentrate on crucial techniques used in developing the suggested models. Section \ref{sec:dl-frameworks} will discuss the deep learning frameworks used in this research. Subsection \ref{sec:data-generators} will explore state-of-the-art models for data generation that are unrelated to audio but can serve as references, such as image generators, to provide further context.

\input{src/chapters/2-state-of-the-art/background/sound.tex}
\input{src/chapters/2-state-of-the-art/background/deep-learning-models.tex}
\input{src/chapters/2-state-of-the-art/background/more.tex}
\input{src/chapters/2-state-of-the-art/background/dl-frameworks}
\input{src/chapters/2-state-of-the-art/background/data-generators.tex}